Reduced growth due to belowground sink limitation is not fully explained by reduced photosynthesis

### 2016-07-08 manuscript/manuscript_pve.Rmd
bm edits to manuscript

In the absence of a good understanding of the effects of sink limitation on physiological processes, most attempts to model sink limitation of growth to date have done so by abandoning the principle of mass balance *(e.g. Leuzinger; CLM nutrient limitation; Fatichi)*. This approach is unsatisfactory because mechanistic models should obey the laws of physics. Our goal in this paper was to investigate how the carbon balance of plants is modified by sink limitation, by experimentally imposing a sink limitation to growth.

*also deleted below 2 sentences*
As plant growth cannot be simply determined by *A*~n~, growth models must take into account several simultaneous processes affecting the C balance at any given time [@fourcaud2008plant]. Whether growth limitation in woody species is driven mostly by *A*~n~ or other components of the C balance remains largely uncertain.

\
In previous studies, the sink limitation of growth has been explored by manipulations of both C source and C sink activity. Examples of experimental approaches that manipulate C source activity include elevated atmospheric CO2 concentration and partial defoliation experiments. Elevated CO~2~ causes an increase in leaf-level photosynthetic rate, *A*~n~ [@drake1997more; @ainsworth2007response]. While this increase is often observed to drive an increase in plant biomass production (eg any number of papers),

across four forest Free-Air CO~2~ Enrichment (FACE) experiments this increased photosynthetic uptake resulted in a stimulation of biomass production by 23 % on average [@norby2005forest]. Evidence from a wide range of elevated CO~2~ experiments, however, also reveals that even with an average photosynthetic enhancement of over 30 %, the biomass growth rate only increases by around 10 % [@kirschbaum2011does].

\
In partial defoliation experiments, source activity is decreased by reduction of photosynthesizing leaf area. In these experiments, compensatory increases in *A*~n~ of the remaining foliage are commonly found, and have been attributed to several processes including indirect effects on resource availability, such as increaed stomatal conductance [@layne1995end] or enhanced leaf nutrient status [@turnbull2007increased; @pinkard2011gas], or more direct effects of modified sink-source balance such as, reduction in end-product inhibition [@iglesias2002regulation; @zhou2003changes; @handa2005test], enhanced biochemical activity [@ovaska1993b;@layne1995end;@pinkard2011gas], and regulatory sugar signaling [@eyles2013whole]. However, there have been few attempts *(any??)* to determine the overall C balance in these experiments, and it is unclear whether the change in growth following defoliation can be explained by the reduction in leaf area and changes to A~n~

Similar to elevated CO~2~ experiments, these increases in *A*~n~ in defoliation experiments did not always produce increased growth due to reductions in meristem sink strength [@palacio2012fast], C limitation to mycorrhizal colonization [@markkola2004defoliation], or an overall decrease in whole plant C gain [@ovaska1993a]. These manipulations of C source activity demonstrate that changes in *A*~n~ do not always correspond with proportional responses of growth.

\
Alternatively, sink limitation can be investigated by manipulating plant tissue C sinks, for example by fruit remove or phloem girdling. In these studies, down regulation of *A*~n~ is often observed, and typically correlates with carbohydrate accumulation resulting from reduced tissue sink strength  [@iglesias2002regulation;@urban2007girdling; @haouari2013fruit]. However, reductions in *A*~n~ have also been attributed to biochemical limitations prior to carbohydrate accumulation [@nebauer2011photosynthesis], irreversible photo-oxidative damage [@duan2008photosynthetic] and stomatal limitation [@li2005photosynthesis]. 
*In addition, it is unclear whether these impacts on An can fully explain the resulting impacts on growth, (? Has anyone attempted to determine this? Is there any evidence for impacts on respiration??)*

\
Yet another experimental approach is to reduce belowground C sink strength in tree seedlings by manipulating container size and thus rooting volume [@arp1991effects; @nesmith1998effect; @poorter2012pot]. Inadequate rooting volume can decrease C sink strength by progressively restricting root growth in growing plants [@thomas1991root]. In a comprehensive meta-analysis, Poorter et al. [-@poorter2012pot] demonstrated a large effect of container size on plant biomass accumulation. Photosynthetic down-regulation is generally observed in these studies [e.g. @arp1991effects; @mcconnaughay1991physical; @gunderson1994photosynthetic; @sage1994acclimation; @maina2002intra; @ronchi2006growth]. Poorter et al. [-@poorter2012pot] examined the components of relative growth rate and found the unit leaf rate to be most strongly affected. As unit leaf rate is correlated with *A*~n~, they concluded that down-regulation of *A*~n~ may best explain the effects of container size on biomass growth. However, they were unable to determine the physiological mechanism  driving the downregulation of *A*~n~, nor to test whether the reduction in *A*~n~ could explain the observed reduction in growth.

@@

\
Decreases in *A*~sat~ occurred simultaneously with reductions of growth of seedlings in containers. This observation initially suggests that sink limitation may be driving the down regulation of photosynthesis, and thus growth. However, there are several possible mechanisms that can explain reduced *A*~n~ in small containers including reduced water availability or nutrient content, increased soil temperatures, or reduced sink strength [@poorter2012pot]. It was therefore necessary to examine each of these factors to determine the mechanism by which the induced belowground sink limitation triggered photosynthetic down regulation. In this study, the impacts of increasing soil temperatures with decreasing container sizes was controlled by burying of all containers flush to the soil surface.

@@

\
As both rooting space and resources were finite in containers, the inability of seedlings to maintain the capacity of the belowground C sink resulted in the buildup of C assimilate in leaves. The feedback inhibition of *A*~n~ from starch accumulation has been proposed, yet it is still not known whether there is a starch threshold that triggers the down-regulation process [@nebauer2011photosynthesis]. Here, declines in *A*~max~ were correlated with higher starch content throughout the experiment. These results agree with a study on a deciduous conifer by Equiza et al. [-@equiza2006photosynthetic] where photosynthetic downregulation from reduced sink strength was correlated with starch content. As starch content in leaves of plants grown in the smallest containers was nearly double that of free seedlings in our study, this suggests the response of *A*~n~ to sink inhibition was regulated by this accumulation, as hypothesized.

##Biomass partitioning under sink limitation
As biomass partitioning is likely controlled by the source and sink strength of all organs [@poorter2012biomass], it was important to determine which tissue components were most affected by the container size treatments. It was necessary to distinguish if growth was affected beyond ontogenetic constraints, by correcting for size, as biomass distribution is strongly size-dependent [@gould1966allometry; @lleonart2000removing]. In this study, there was no significant difference in root, leaf, or stem biomass partitioning with reduced soil volume compared to free seedlings, once ontogenetic drift was taken into account (Figure `r allocfig`a,b). This result is unexpected as shifts in partitioning have been noted specifically for nutrient limitation [@mcconnaughay1999biomass, and references therein]. Surprisingly, a constant ratio of fine root mass to leaf mass was observed across all treatments suggesting a functional partitioning response to optimize resource gain did not occur.

@@

##Conclusions
With a novel field-based design we detected a massive effect of container volume on seedling growth, when compared with naturally planted seedlings. This finding is important as manipulations of plants grown in containers are often used to draw conclusions about growth and physiological principles, but how these results actually reflect field-grown plants is seldom studied. Although biomass partitioning was conserved, our empirical and model results suggest that the amount of photosynthate allocated to non-biomass pools were likely altered by sink inhibition. The debate over how rates of photosynthesis affect plant growth or to what degree these rates are instead controlled by growth has existed for decades [@sweet1966role]. Our combined empirical and modeling approach shows that when non-photosynthetic parameters were kept constant, changes in *A*~n~ were not able to fully to predict changes in growth, an important distinction often missed in studies that manipulate source/sink activity. It has been suggested that tissue sink activity strongly feedbacks onto source activity, causing growth to control *A*~n~ through the demand for C [@korner2013growth; @korner2015paradigm]. Although this may be true, our results imply that quantifying the fate of assimilated C into known pools of growth, storage and C loss are needed prior to addressing this debate. The approach used here has the flexibility to account for multiple drivers of C allocation and provides an avenue to address future questions regarding the impact of environmental change on plant growth.

>>>>

In the absence of a good understanding of the effects of sink limitation on physiological processes, most attempts to model sink limitation of growth to date have done so by abandoning the principle of mass balance [e.g. @fatichi2013reconciling; @fatichi2014moving]. This approach is unsatisfactory because mechanistic models should obey the laws of physics. Growth models must take into account several simultaneous processes affecting the C balance at any given time [@fourcaud2008plant], thus incomplete modelling approaches often lead to unaccounted for C. Our goal in this paper was to investigate how the carbon balance of plants is modified by sink limitation, by experimentally imposing a sink limitation to growth.

\
In previous studies, the sink limitation of growth has been explored by manipulations of both C source and C sink activity. Examples of experimental approaches that manipulate C source activity include elevated atmospheric CO2 concentration and partial defoliation experiments. Elevated CO~2~ causes an increase in leaf-level photosynthetic rate, *A*~n~ [@drake1997more; @ainsworth2007response]. This increased photosynthetic uptake is often observed to drive an increase in plant biomass production [@norby2005forest]. Evidence from a wide range of elevated CO~2~ experiments, however, also reveals that even with an average photosynthetic enhancement of over 30 %, the biomass growth rate only increases by around 10 % [@kirschbaum2011does]. This implies that other process of tree C balance must be affected, which has been noted in CO~2~ manipulation studies that attempt to close the C budget [@wang1998carbon; @hamilton2002forest; @schafer2003exposure].

\
In partial defoliation experiments, source activity is decreased by reduction of photosynthesizing leaf area. In these experiments, compensatory increases in *A*~n~ of the remaining foliage are commonly found, and have been attributed to several processes including indirect effects on resource availability, such as increaed stomatal conductance [@layne1995end] or enhanced leaf nutrient status [@turnbull2007increased; @pinkard2011gas], or more direct effects of modified sink-source balance such as, reduction in end-product inhibition [@iglesias2002regulation; @zhou2003changes; @handa2005test], enhanced biochemical activity [@ovaska1993b;@layne1995end;@pinkard2011gas], and regulatory sugar signaling [@eyles2013whole]. Similar to elevated CO~2~ experiments, these increases in *A*~n~ in defoliation experiments do not increase tree growth [@palacio2012fast; @markkola2004defoliation; @ovaska1993a]. As these studies do not attempt to scale the degree of compensentory response in leaf C gain to the entire canopy, the overall affect of source manipulation on tree C balance and growth is still unclear.

\
Alternatively, sink limitation can be investigated by manipulating plant tissue C sinks, for example by fruit remove or phloem girdling. In these studies, down regulation of *A*~n~ is often observed, and typically correlates with carbohydrate accumulation resulting from reduced tissue sink strength  [@iglesias2002regulation;@urban2007girdling; @haouari2013fruit]. However, reductions in *A*~n~ have also been attributed to biochemical limitations prior to carbohydrate accumulation [@nebauer2011photosynthesis], irreversible photo-oxidative damage [@duan2008photosynthetic] and stomatal limitation [@li2005photosynthesis]. 
Similar to C source manipulations, it is still unclear whether these impacts on *A*~n~ can fully explain the resulting impacts on growth.

\
Yet another experimental approach is to reduce belowground C sink strength in tree seedlings by manipulating container size and thus rooting volume [@arp1991effects; @nesmith1998effect; @poorter2012pot]. Inadequate rooting volume can decrease C sink strength by progressively restricting root growth in growing plants [@thomas1991root]. In a comprehensive meta-analysis, Poorter et al. [-@poorter2012pot] demonstrated a large effect of container size on plant biomass accumulation. Photosynthetic down-regulation is generally observed in these studies [e.g. @arp1991effects; @mcconnaughay1991physical; @gunderson1994photosynthetic; @sage1994acclimation; @maina2002intra; @ronchi2006growth]. Poorter et al. [-@poorter2012pot] examined the components of relative growth rate and found the unit leaf rate to be most strongly affected by rooting volume. As unit leaf rate is correlated with *A*~n~, they concluded that down-regulation of *A*~n~ may best explain the effects of container size on biomass growth. However, they were unable to determine the physiological mechanism  driving the downregulation of *A*~n~, nor to test whether the reduction in *A*~n~ could explain the observed reduction in growth.

@@

\
Decreases in *A*~sat~ occurred simultaneously with reductions of growth of seedlings in containers. This observation initially suggests that sink limitation may be driving the down regulation of photosynthesis, and thus growth. However, there are several possible mechanisms that can explain reduced *A*~n~ in small containers including increased soil temperatures, reduced water availability or nutrient content, or reduced sink strength [@poorter2012pot]. It was therefore necessary to examine each of these factors to determine the mechanism by which the induced belowground sink limitation triggered photosynthetic down regulation. In this study, the impacts of increasing soil temperatures with decreasing container sizes was controlled by burying of all containers flush to the soil surface.

@@

\
As both rooting space and resources were finite in containers, the inability of seedlings to maintain the capacity of the belowground C sink resulted in the buildup of C assimilate in leaves. The feedback inhibition of *A*~n~ from starch accumulation has been proposed, yet it is still not known whether there is a starch threshold that triggers the down-regulation process [@nebauer2011photosynthesis]. Here, declines in *A*~max~ were correlated with higher starch content throughout the experiment. These results agree with a study on a deciduous conifer by Equiza et al. [-@equiza2006photosynthetic] where photosynthetic downregulation from reduced sink strength was correlated with starch content. As starch content in leaves of plants grown in the smallest containers was nearly double that of free seedlings in our study, this suggests the response of *A*~n~ to sink inhibition was regulated by this accumulation, as hypothesized. Overall, it is likely that both nutrient content and reduced sink strength played a role in observed photosynthetic down regulation, but future studies are needed to identify the specific mechanisms which prommpted these feedbacks.

##Biomass partitioning under sink limitation
As biomass partitioning is likely controlled by the source and sink strength of all organs [@poorter2012biomass], it was important to determine which tissue components were most affected by the container size treatments. It was necessary to distinguish if growth was affected beyond ontogenetic constraints, by correcting for size, as biomass distribution is strongly size-dependent [@gould1966allometry; @lleonart2000removing]. In this study, there was no significant difference in root, leaf, or stem biomass partitioning with reduced soil volume compared to free seedlings, once ontogenetic drift was taken into account (Figure `r allocfig`a,b). Although shifts in partitioning have been previosuly reported, specifically for nutrient limitation [@mcconnaughay1999biomass, and references therein], a constant ratio of fine root mass to leaf mass was observed in this study across all treatments. These results suggest that seedlings kept a conservative strategy to maintain homeostatis with with biomass partitioning, instead of a functional partitioning response to optimize limiting soil resources.

@@

##Conclusions
With a novel field-based design we detected a massive effect of container volume on seedling growth, when compared with naturally planted seedlings. This finding is important as manipulations of plants grown in containers are often used to draw conclusions about growth and physiological principles, but how these results actually reflect field-grown plants is seldom studied. As seedlings maintained a conservative partitioning strategy, aboveground biomass was restricted in coordination with root restriction from limited soil volume. Photosynthesis was affected by both reduced nutrient uptake and the buildup of starch, both potential mechanism for downregulation of *A*~n~. Our combined empirical and modeling approach shows that when non-photosynthetic parameters were kept constant, changes in *A*~n~ were not able to fully to predict changes in growth, an important distinction often missed in studies that manipulate source/sink activity. It has been suggested that tissue sink activity strongly feedbacks onto source activity, causing growth to control *A*~n~ through the demand for C [@korner2013growth; @korner2015paradigm]. Our results imply that testing this hypothesis requires accurate modelling of C mass balance before addressing this debate.
